Dish: Distribution-oblivious Shell Scripting

### 2020-02-29 pldi/draft.tex
Rewrite and refine of 4.1

The core of \sys is a dataflow graph model, inspired by the ones used
in popular distributed stream processing systems~\sx{related}. The
dataflow model can express a large subset of Shell programs and
inherently supports the data parallelism found in shell pipelines by
exposing it as task parallelism. More preciely, we have developed a
set of parallelization-exposing optimizations represented as
semantics-preserving graph transformations. These transformations can
then be applied in an iterative way, optimizing the program to utilize
available computational resources.

@@

\kk{It might be beneficial to do a full pass over this section
  describing the new model (where a node that has many sequential
  inputs is always prepended with a cat node.}

@@

@@

For a set $D$, we write $D\kstar$ to denote the set of all finite
words over $D$. For words $x, y \in D\kstar$, we write $x \cdot y$ or
$xy$ to denote their concatenation. We write $\eps$ for the empty
word. We say that $x$ is a \emph{prefix} of $y$, and we write $x \leq
y$, if there is a word $z$ such that $y = xz$. It is easy to see that
$\leq$ is a partial order, often called the \emph{prefix order}.

A node $f$ of the graph represents a function from one (possibly
empty) input stream to an output stream $f : D\kstar \to D\kstar$,
where $D$ represents the basic data type of a line of characters
\kk{We have to extend this to more general nodes. It has to be from
  many inputs to many outputs. The commands that we optimize at the
  moment are the ones that can be rewritten to take one stream input
  to one stream output, and then some other inputs and outputs that
  are considered static.}. This representation captures the majority
of shell commands. We require that the function $f$ is monotone
w.r.t.\ the prefix order. This captures the idea that a node cannot
retract output after it has been emitted.
%More formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there
%exists $y'$ such that $f(x.x') = y.y'$, where $.$ is standard
%concatenation and will be omitted when obvious.
An example of a node is the following command:

@@

An important
observation is that commands might read their input
stream from a sequence of different sources. An illustrative example
is the command \ttt{cat}, which reads its input from a combination
of files and its standard input:

@@

\noindent
In the example above, \ttt{cat} reads file \ttt{x} until it encounters the EOF
character, then reads from \ttt{stdin} until in encounters EOF, and finally reads from
\ttt{y}.

In order to handle this particularity, nodes in the
dataflow graph can have many incoming edges, that are \emph{ordered} in a
sequence. This is different from standard dataflow
graph representations where the semantics of multiple incoming edges
is either zipping the inputs into pairs, or interleaving them arbitrarily.

%% \TODO{Is the following paragraph best suited here? It describes an
%%   assumption that we make about commands, to guarantee the validity of
%%   our transformations. The assumption is needed due to the limitation
%%   of the programming model of one input one output per command.}

Nodes may read from more files than just the files of their
input stream (resp. write to more files than their output stream).
However, these files are static (\eg used for configuration or
logging) and do not represent streams of data, and therefore are not
considered part of the dataflow graph.
For now, we assume that these static
files can only be accessed by one node, thus ensuring that they do not
interfere with the distributed implementation.
%% Because of the assumption in section Front End \ref{}, that each file
%% exists only once in the dataflow graph, we can safely assume that
%% reading and writing to these static files does not interfere with the
%% distributed implementation, and doesn't alter the behaviour of the
%% program.

Edges in the dataflow graph represent files, the basic data
abstraction of the shell. They are used as communication channels
between nodes in the graph, and as the input or output of the whole
graph. Edges are represented as possibly unbounded streams of type
$D\kstar$. They could either refer to a specific named file in the file
system, or just be FIFO pipes used for interprocess communication
%% \TODO{Are there any other file types, such as URLs, ...?}.

In order to expose data parallelism from the output of a command, files that
do not point to a resource can have
\begin{wrapfigure}[3]{r}{0.20\columnwidth}
\vspace{-7pt}
  \includegraphics[width=0.20\columnwidth]{\detokenize{./figs/dish_g2.pdf}}
\vspace{-20pt}
\end{wrapfigure}
an upper bound on the lines that they can transfer. This helps utilize available
parallelism after encountering a command that cannot be parallelized
(\eg \ttt{sha1sum}). The schematic on the right illustrates this scenario.

%% \kk{(Maybe) It might be beneficial to mention that many inputs do not
%%   represent the same thing as many outputs. These two notions are not
%%   symmetric.}

@@

@@

@@

>>>>

The core of \sys is a dataflow graph model, inspired by the ones used
in popular distributed stream processing systems~\sx{related} \kk{Is
  this reference here necessary?}. The dataflow model can express a
large subset of Shell programs and inherently supports the data
parallelism found in shell pipelines by exposing it as task
parallelism. More preciely, we have developed a set of
parallelization-exposing optimizations represented as
semantics-preserving graph transformations. These transformations can
then be applied in an iterative way, optimizing the program to utilize
available computational resources.

@@

@@

Before describing the graph edges, we introduce basic notation that
will be useful later on.  For a set $D$, we write $D\kstar$ to denote
the set of all finite words over $D$. For words $x, y \in D\kstar$, we
write $x \cdot y$ or $xy$ to denote their concatenation. We write
$\eps$ for the empty word. We say that $x$ is a \emph{prefix} of $y$,
and we write $x \leq y$, if there is a word $z$ such that $y = xz$. It
is easy to see that $\leq$ is a partial order, often called the
\emph{prefix order}.

Edges in the dataflow graph represent files, the basic data
abstraction of the shell. They are used as communication channels
between nodes in the graph, and as the input or output of the whole
graph. Edges are represented as possibly unbounded streams of type
$D\kstar$. They could either refer to a specific named file in the file
system, or just be FIFO pipes used for interprocess communication.

%% \kk{I think we should delete this since the node and not the edge
%%   chooses how to write to its output files.}

%% In order to expose data parallelism from the output of a command,
%% files that do not point to a resource can have
%% \begin{wrapfigure}[3]{r}{0.20\columnwidth}
%% \vspace{-7pt}
%%   \includegraphics[width=0.20\columnwidth]{\detokenize{./figs/dish_g2.pdf}}
%% \vspace{-20pt}
%% \end{wrapfigure}
%% an upper bound on the lines that they can transfer. This helps utilize available
%% parallelism after encountering a command that cannot be parallelized
%% (\eg \ttt{sha1sum}). The schematic on the right illustrates this scenario.

@@

A node $f$ of the graph represents a function from a (possibly empty)
list of input streams to a list of output streams $f : [D\kstar] \to
[D\kstar]$, where $D$ represents the basic data type of a line of
characters. This representation captures the majority of shell
commands. We require that the function $f$ is monotone w.r.t.\ a
lifting of the prefix order for a sequence of inputs \kk{Formalize
this lifting if essential for later development}. This captures the
  idea that a node cannot retract output after it has been emitted.
%More formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there
%exists $y'$ such that $f(x.x') = y.y'$, where $.$ is standard
%concatenation and will be omitted when obvious.
An example of a node with one input and one output stream is the
following command:

@@

An important observation is that even though commands might read their
input from several input streams, the order in which they read is not
always arbitrary. An illustrative example is the command \ttt{cat},
which consumes its input one stream at a time:

@@

\noindent
In the example above, \ttt{cat} reads file \ttt{x} until it encounters
the EOF character, then reads from \ttt{stdin} until in encounters
EOF, and finally reads from \ttt{y}. Other commands that read their
inputs one file at a time include \ttt{grep} and \ttt{tr}. Knowing the
order in which a node consumes its input is necessary to enable
certain optimizations on the graph.

%% \kk{Some could read in sequence (like cat), some in lockstep (like
%%   zip), some read some files statically and the rest as streams (like
%%   comm), some read arbitrarily.}

\kk{I need to mention that there are nodes that we know more
  properties of. For example cat, tee, zip, etc... and these are
  handled specially because of their very important and special
  properties.}

@@

\kk{I have to mention that in order for the transformations to happen,
  we transform all commands that read some of their inputs in sequence
  to commands that read from one input but are preceded by a cat.}

\kk{I have to mention that all stateless and pure commands fall into
  the category of commands that have one input stream (and possibly
  several input static files) and one output stream (and no(?!?)
  output static files?). These commands are the bulk of
  data-processing in the shell, and these are the ones that we handle
  in our system.}

\kk{Move this assumption in the transformation}
Nodes may read from more files than just the files of their
input stream (resp. write to more files than their output stream).
However, these files are static (\eg used for configuration or
logging) and do not represent streams of data, and therefore are not
considered part of the dataflow graph.
For now, we assume that these static
files can only be accessed by one node, thus ensuring that they do not
interfere with the distributed implementation.
%% Because of the assumption in section Front End \ref{}, that each file
%% exists only once in the dataflow graph, we can safely assume that
%% reading and writing to these static files does not interfere with the
%% distributed implementation, and doesn't alter the behaviour of the
%% program.

@@

\kk{We should move FIgure 4 here}

@@

\kk{This subsection has to only talk about the cool shell constructs
  that are added and about optimizations that are not in the dataflow
  graph (like file splitting), and the implementation of split, the
  pure merges , etc...}
